---
title       : NHL Team Points Predictor
subtitle    : 2014-15 Season
author      : David J Freifeld
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Data

The official website of the National Hockey League, nhl.com, provides basic data for all 30 teams for each season. For the current, 2014-15 season, we take a subset of this data, including the following five variables:

1. Goals For
2. Goals Against
3. Power Play Percentage
4. Penalty Kill Percentage
5. Points

--- .class #id 

## Motivation

- A team's position in the playoffs is dependent on the number of points they accumulate throughout the season. 
- A team receives 2 points for a win.
- A team receives 1 point if they lose in overtime. 
- A team or a fan could use this tool to set goals for how many goals they want to score, how few goals to allow, etc., to get to a certain point total.
- A linear multivariate regression is ideal because it is quick, simple, and easily interpretable.

---

## Points vs. Goals For and Goals Against

```{r warning=FALSE, echo=FALSE}
library(ggplot2)
nhl <- read.csv("NHL14-15.csv")
```
```{r}
print(ggplot(nhl, aes(x=GF, y=GA, color=P)) + geom_point(size=4))
```

---

## Takeaways

- As we saw in the previous plot, a team is more likely to score more points when it scores more goals and has fewer goals scored against it.
- Interestingly, though, the coefficient of the Power Play Percentage variable is negative when using all four covariates:

```{r}
summary(lm(P ~ ., data=nhl))$coefficients
```

- This could be because, when holding the number of Goals a team scored constant, a higher power play percentage may mean that a team is too dependent on power plays to score. 
- Regardless, there are clearly some interesting insights that can be gained from this tool.


